---
title: "Project 2"
author: "Team 8 - C/D"
date: "4/26/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Data & Libraries
```{r load, message: FALSE}
library(quantmod)
library(forecast)
library(timeSeries)
library(corrplot)

rm(list = ls())
data1 <- read.csv("BPPTrain.csv", header = TRUE)[,c(1,4)]
data2 <- read.csv("BPPTrain.csv", header = TRUE)[,c(1,2,3)]

colnames(data1)[1] <- "date"
colnames(data2)[1] <- "date"

data1[data1=="#N/A"] = NA #change #N/A to NA
data1$NASDAQ = as.numeric(data1$NASDAQ) 
data <- as.zoo(merge(as.timeSeries(data1), as.timeSeries(data2)))

data = data[complete.cases(data),]
```

## Data Preparation
```{r data prep}
#calculate day-to-day difference
dnas   <- diff(data$NASDAQ, na.pad = TRUE)
dbpp   <- diff(data$BPP, na.pad = TRUE)
dcpi   <- diff(data$CPI, na.pad = TRUE)
#lag 1 row
NASDAQ1     <- lag(data$NASDAQ, na.pad = TRUE, k=-1)
dnas1       <- lag(dnas, na.pad = TRUE, k=-1)
dbpp1       <- lag(dbpp, na.pad = TRUE, k=-1)
dcpi1       <- lag(dcpi, na.pad = TRUE, k=-1)
dnas2       <- lag(dnas, na.pad = TRUE, k =-2)
dbpp2       <- lag(dbpp, na.pad = TRUE, k=-2)
dcpi2       <- lag(dcpi, na.pad = TRUE, k=-2)
nas_3d_mean <- lag(rollapply(data$NASDAQ, 3, mean,fill=NA,align = "right"),k=-1)
nas_3d_max  <- lag(rollapply(data$NASDAQ, 3, max,fill=NA,align = "right"),k=-1)
bpp_3d_median <- lag(rollapply(data$BPP, 3,median,fill=NA,align = "right"),k=-1)
bpp_3d_max  <- lag(rollapply(data$BPP, 3,max,fill=NA,align = "right"),k=-1)
nas_10d_mean <- lag(rollapply(data$NASDAQ, 10, mean,fill=NA,align = "right"),k=-1)
nas_20d_mean <- lag(rollapply(data$NASDAQ, 20, mean,fill=NA,align = "right"),k=-1)

cumavg.nas  <- lag(cumsum(c(data$NASDAQ))/seq(1:nrow(data)),na.pad=TRUE,k=-1)
#calculate actual return y
return      <- data$NASDAQ/NASDAQ1
#add calculated fields back to data
data       <- cbind(data, dnas, dbpp, dcpi, 
                    dnas1, dbpp1, dcpi1, dnas2, dbpp2, dcpi2, 
                    nas_3d_mean,nas_3d_max,bpp_3d_median,bpp_3d_max,nas_10d_mean,nas_20d_mean,
                    cumavg.nas,return)
```

```{r}
library(dplyr)

#create time series split
sequence <- seq(from=1, to=nrow(data))
in.sample <- sequence<1500
out.sample = nrow(data)-sum(in.sample)

data.in  <- data[in.sample,] #training set
data.val <- data[!in.sample,] #validation set

#split validation to val & test
sequence <- seq(from=1, to=out.sample)
validation <- sequence<500
data.val2  <- data.val[validation,] #training set
data.test <- data.val[!validation,] #validation set

```


## Data Exploration/Visualization
```{r}
plot(data$NASDAQ)
plot(dnas)
plot(dbpp1,dnas)
plot(dcpi1,dnas)
plot(bpp_3d_median,dnas)
```

```{r}
(cor(data.val, data.val$dnas))
```

## Investment Strategies

### Strategy 1 - get best quantile

We cross-validate on the validation set to find the best range to invest, the boundaries are calculated according to quantiles of the predicted values on validation set.

```{r get_quantile function}
get_quantile <- function(pred,val){
  prob_range1 = seq(0, 0.5, by = 0.01)
  prob_range2 = seq(1, 0.5, by = -0.01)
  IR_best = 0
  i_best = 0
  j_best = 0
  return_best = 0
  for(i in prob_range1){
    for (j in prob_range2){
      buy.out   <- c(as.numeric(pred>=quantile(pred, probs = i)&
                                pred<=quantile(pred, probs = j)))
      dailyreturn.out  <- (val$return-1)*buy.out
      #print(dailyreturn.out)
      IR  <- mean(dailyreturn.out)/sd(dailyreturn.out)
      #print(IR)
      if(is.nan(IR)){
        IR = 0
        }
      if(IR>IR_best){
        IR_best = IR
        i_best = quantile(pred, probs = i)
        j_best = quantile(pred, probs = j)
        return_best = dailyreturn.out
      }
    }
  }
  return(c(IR_best,i_best,j_best,return_best))
}
```

### Strategy 2 - get best threshold

This method is a correction of best_quantile method. The algorithm decides the best predicted `dnas` threshold for investment to be made based on validation set IR (pick the highest one). This method can be applied as a real trading strategy.

```{r find best threshold}
get_thres <- function(pred,val){
  thres_range = seq(-20, 10, by = 0.1)
  IR_best = 0
  i_best = 0
  return_best = 0
  IR_range = replicate(length(thres_range),0)
  for(j in 1:length(thres_range)){
    i = thres_range[j]
    buy.out   <- c(as.numeric(pred>=i))
    dailyreturn.out  <- (val$return-1)*buy.out
    IR  <- mean(dailyreturn.out)/sd(dailyreturn.out)
    IR_range[j] = IR
    if(IR>=IR_best & !is.na(IR)){
      IR_best = IR
      i_best = i
      return_best = dailyreturn.out
    }
  }
  plot(thres_range,IR_range,bg = 'gray',xlab="threshold",ylab="IR")
  text(x=thres_range[IR_range==max(IR_range)], y=max(IR_range), labels="best threshold",col="red")
  return(c(IR_best,i_best,return_best))
}
```

### Strategy 3 - rolling average method

This method calculates the rolling average of previous 5,10,20,30,50,100 days, and find the best n day that we should calculate for the predicted validation set. If the predicted `NASDAQ` value falls below the n-day mean minus 1 standard deviation, we buy, and if it falls above the n-day mean plus 1 standard deviation, we sell.

```{r rolling_method}
rolling_method <- function(pred,val){
  n = 0
  bestIR = 0
  return = replicate(length(pred), 0)
  n_range = c(5,10,20,30,50,100)
  for (j in n_range){
    alldata = rbind(data.in$return,pred)
    rolling_mean <- rollapply(alldata, j, mean,fill=NA,align = "right")
    rolling_mean <- rolling_mean[-(1:sum(in.sample))]
    rolling_sd <- rollapply(alldata, j, sd,fill=NA,align = "right")
    rolling_sd <- rolling_sd[-(1:sum(in.sample))]
    buy.out <- replicate(length(pred), 0)
    status = "out"
    for (i in 1:length(pred)){
      if(pred[i]<=rolling_mean[i]-rolling_sd[i]){
        buy.out[i] = 1
        status = "in"
      }else if(pred[i]>=rolling_mean[i]+rolling_sd[i] & status == "in"){
        buy.out[i] = 1
        status = "out"
      }else{
        if(status == "in"){
          buy.out[i]=1
        }
      }
    }
    dailyreturn.out  <- (val$return-1)*buy.out
    IR  <- mean(dailyreturn.out)/sd(dailyreturn.out)
    if(IR >bestIR){
      bestIR = IR
      n = j
      return = dailyreturn.out
    }
  }
  return(c(bestIR,n,return))
}
```

### Preliminary evaluations

All-in method:
For simplicity, we assume every time we will invest all money into NASDAQ and sell all holdings at the end of the day.

```{r}
all_in<-function(return){
  
  cashovertime = 1:length(return)+1
  cash.out = 10000000
  cashovertime[1] = cash.out
  #calculated compounded return
  for(i in 1:length(return)){
      cash.out <- cash.out*(as.numeric((return[i]+1)))
      cashovertime[i+1] = cash.out
  }
  result = list()
  result$cashreturn = (cash.out - 10000000)/10000000
  result$overtime = cashovertime
  return(result)
}
```


Let's first see what the baseline is:

```{r baseline}
#calculate the baseline: holding NASDAQ all the time
buy.out_baseline <- replicate(out.sample,1) #baseline (holding NASDAQ)
dailyreturn.out_baseline  <- (data.val$return-1)*buy.out_baseline
IR_baseline  <- mean(dailyreturn.out_baseline)/sd(dailyreturn.out_baseline)
mean_return_baseline = mean(dailyreturn.out_baseline)
cash_baseline = all_in(dailyreturn.out_baseline)
cash_return_baseline = cash_baseline$cashreturn
cash_overtime_baseline = cash_baseline$overtime
data.frame(IR = IR_baseline,mean_return = mean_return_baseline,cash_return = cash_return_baseline)
```

use the simplest model to pick the best method
```{r}
#use data.val2, which is a subset of validation set, and test of data.test which is a hold-out
lM1     <- lm(dnas ~ dnas1, data=data.in)
pred.outl1   <- predict(lM1,  newdata=data.val2)

lM1nasdaq     <- lm(NASDAQ ~ dnas1, data=data.in)
pred.outl1nasdaq   <- predict(lM1nasdaq,  newdata=data.val2)

l1_q  <- get_quantile(pred.outl1,data.val2)
l1_lower = l1_q[2]
l1_upper = l1_q[3]
l1_t  <- get_thres(pred.outl1,data.val2)
l1_thres = l1_t[2]
l1_r  <- rolling_method(pred.outl1nasdaq,data.val2)

pred.outl1_test   <- predict(lM1,  newdata=data.test)
buy.out_q   <- c(as.numeric(pred.outl1_test>=l1_lower&
                                pred.outl1_test<=l1_upper))
dailyreturn.out_q  <- (data.test$return-1)*buy.out_q
IR_q  <- mean(dailyreturn.out_q)/sd(dailyreturn.out_q)
      
buy.out_t   <- c(as.numeric(pred.outl1_test>=l1_thres))
dailyreturn.out_t  <- (data.test$return-1)*buy.out_t
IR_t  <- mean(dailyreturn.out_t)/sd(dailyreturn.out_t)

data.frame(quantile_method=IR_q,
           threshold_method = IR_t,
           rolling_method = l1_r[1])
```
According to our preliminary comparison, we found out that using the best_quantile method works best. Our get_threshold methods also outputs a plot so we can seee how the IR changes with threshold.

Let's compare it with the baseline on whole validation:
```{r}
pred.outl1   <- predict(lM1,  newdata=data.val)
buy.out_q   <- c(as.numeric(pred.outl1>=l1_lower&
                            pred.outl1<=l1_upper))
dailyreturn.out_q  <- (data.val$return-1)*buy.out_q
IR_q  <- mean(dailyreturn.out_q)/sd(dailyreturn.out_q)
      
buy.out_t   <- c(as.numeric(pred.outl1>=l1_thres))
dailyreturn.out_t  <- (data.val$return-1)*buy.out_t
IR_t  <- mean(dailyreturn.out_t)/sd(dailyreturn.out_t)
l1_r  <- rolling_method(pred.outl1,data.val)

data.frame(baseline = IR_baseline,
          quantile_method=IR_q,
           threshold_method = IR_t,
           rolling_method = l1_r[1])
```
Quantile method is still the best,but we are going to test both quantile & threshold for future modeling.

## Modeling

### Linear Regressions
```{r linear reg}
#OLS
lM1     <- lm(dnas ~ dnas1, data=data.in)
lM2     <- lm(dnas ~ dnas1 + dbpp1 + dcpi1, data=data.in)
lM3     <- lm(dnas ~ dnas1 + dbpp1 + dcpi1 + dbpp2 + dcpi2,data=data.in)
lM4     <- lm(dnas ~ .-dnas-NASDAQ-return-dbpp-dcpi-CPI-BPP,data=data.in)
lM5     <- lm(dnas ~ dnas1+bpp_3d_median,data=data.in)

#predicted return = y_hat
pred.outl1   <- predict(lM1,  newdata=data.val)
pred.outl2   <- predict(lM2,  newdata=data.val)
pred.outl3   <- predict(lM3, newdata=data.val)
pred.outl4   <- predict(lM4, newdata=data.val)
pred.outl5   <- predict(lM5, newdata=data.val)

#apply investment strategy for each model
lM1_q  <- get_quantile(pred.outl1,data.val)
IR_lM1 = lM1_q[1]
returnl1  <- lM1_q[-(1:3)]
mean_return_lM1 = mean(returnl1)
cash_lM1 = all_in(returnl1)
cash_return_lM1 = cash_lM1$cashreturn
cash_overtime_lM1 = cash_lM1$overtime

lM1_t  <- get_thres(pred.outl1,data.val)
IR_lM1_t = lM1_t[1]
returnl1_t  <- lM1_t[-(1:2)]
mean_return_lM1_t = mean(returnl1_t)
cash_lM1_t = all_in(returnl1_t)
cash_return_lM1_t = cash_lM1_t$cashreturn
cash_overtime_lM1_t = cash_lM1_t$overtime

lM2_q  <- get_quantile(pred.outl2,data.val)
IR_lM2 = lM2_q[1]
returnl2  <- lM2_q[-(1:3)]
mean_return_lM2 = mean(returnl2)
cash_lM2 = all_in(returnl2)
cash_return_lM2 = cash_lM2$cashreturn
cash_overtime_lM2 = cash_lM2$overtime

lM2_t  <- get_thres(pred.outl2,data.val)
IR_lM2_t = lM2_t[1]
returnl2_t  <- lM2_t[-(1:2)]
mean_return_lM2_t = mean(returnl2_t)
cash_lM2_t = all_in(returnl2_t)
cash_return_lM2_t = cash_lM2_t$cashreturn
cash_overtime_lM2_t = cash_lM2_t$overtime

lM3_q  <- get_quantile(pred.outl3,data.val)
IR_lM3 = lM3_q[1]
returnl3  <- lM3_q[-(1:3)]
mean_return_lM3 = mean(returnl3)
cash_lM3 = all_in(returnl3)
cash_return_lM3 = cash_lM3$cashreturn
cash_overtime_lM3 = cash_lM3$overtime

lM3_t  <- get_thres(pred.outl3,data.val)
IR_lM3_t = lM3_t[1]
returnl3_t  <- lM3_t[-(1:2)]
mean_return_lM3_t = mean(returnl3_t)
cash_lM3_t = all_in(returnl3_t)
cash_return_lM3_t = cash_lM3_t$cashreturn
cash_overtime_lM3_t = cash_lM3_t$overtime


lM4_q  <- get_quantile(pred.outl4,data.val)
IR_lM4 = lM4_q[1]
returnl4  <- lM4_q[-(1:3)]
mean_return_lM4 = mean(returnl4)
cash_lM4 = all_in(returnl4)
cash_return_lM4 = cash_lM4$cashreturn
cash_overtime_lM4 = cash_lM4$overtime


lM4_t  <- get_thres(pred.outl4,data.val)
IR_lM4_t = lM4_t[1]
returnl4_t  <- lM4_t[-(1:2)]
mean_return_lM4_t = mean(returnl4_t)
cash_lM4_t = all_in(returnl4_t)
cash_return_lM4_t = cash_lM4_t$cashreturn
cash_overtime_lM4_t = cash_lM4_t$overtime


lM5_q  <- get_quantile(pred.outl5,data.val)
IR_lM5 = lM5_q[1]
returnl5  <- lM5_q[-(1:3)]
mean_return_lM5 = mean(returnl5)
cash_lM5 = all_in(returnl5)
cash_return_lM5 = cash_lM5$cashreturn
cash_overtime_lM5 = cash_lM5$overtime


lM5_t  <- get_thres(pred.outl5,data.val)
IR_lM5_t = lM5_t[1]
returnl5_t  <- lM5_t[-(1:2)]
mean_return_lM5_t = mean(returnl5_t)
cash_lM5_t = all_in(returnl5_t)
cash_return_lM5_t = cash_lM5_t$cashreturn
cash_overtime_lM5_t = cash_lM5_t$overtime


output = data.frame(InformationRatio_q = c(IR_lM1,IR_lM2,IR_lM3,IR_lM4,IR_lM5),
           MeanReturn_q = c(mean_return_lM1,mean_return_lM2,mean_return_lM3,
                            mean_return_lM4,mean_return_lM5),
           CashReturn_q = c(cash_return_lM1,cash_return_lM2,cash_return_lM3,
                            cash_return_lM4,cash_return_lM5),
           InformationRatio_t = c(IR_lM1_t,IR_lM2_t,IR_lM3_t,IR_lM4_t,IR_lM5_t),
           MeanReturn_t = c(mean_return_lM1_t,mean_return_lM2_t,
                            mean_return_lM3_t,mean_return_lM4_t,mean_return_lM5_t),
           CashReturn_t = c(cash_return_lM1_t,cash_return_lM2_t,
                            cash_return_lM3_t,cash_return_lM4_t,cash_return_lM5_t))
row.names(output) <- c("lM1","lM2","lM3","lM4","lM5")
output
```
The best-performing linear model(lM5) has dnas1+bpp_3d_median as the independent variable.

```{r}
plot(cash_overtime_lM5, type='l', col='green', xlab='Days',ylab='Cumulative Cash Value ($)')
lines(cash_overtime_lM1,col='gray')
lines(cash_overtime_lM2,col='gray')
lines(cash_overtime_lM3,col='gray')
lines(cash_overtime_lM4,col='gray')
lines(cash_overtime_lM1_t,col='gray')
lines(cash_overtime_lM2_t,col='gray')
lines(cash_overtime_lM3_t,col='gray')
lines(cash_overtime_lM4_t,col='gray')
lines(cash_overtime_lM5_t,col='gray')
lines(cash_overtime_baseline,col='black')
lines(cash_overtime_lM5, col='green', lwd=2)
title("Cumulative Cash Value over Time, OLS Models, out-of-sample")
legend(1,1.4e7,legend=c("LM5 Quantile", "Baseline"), lty=1,col=c("green", "black"))
```


### Lasso

```{r lasso,message = FALSE}
set.seed(1)
library(tidyverse)
library(caret)
library(glmnet)
# Predictor variables
excl = c('dnas','NASDAQ','CPI','BPP','return','dbpp','dcpi')
x.train <- model.matrix(dnas~., data.in)
x.train = x.train[,!(colnames(x.train) %in% excl)]
x.val <- model.matrix(dnas~., data.val)
x.val = x.val[,!(colnames(x.val) %in% excl)]
# Outcome variable
diff = nrow(data.in) - nrow(x.train)
y.train <- data.in$dnas[-(1:diff)]
cv <- cv.glmnet(x.train, y.train, alpha = 1)

lasso <- glmnet(x.train, y.train, alpha = 1, lambda = cv$lambda.min)

#predicted return = y_hat
pred.outlasso <- predict(lasso,newx=x.val)

lasso_q  <- get_quantile(pred.outlasso,data.val)
IR_lasso = lasso_q[1]
returnlasso <- lasso_q[-(1:3)]
mean_return_lasso = mean(returnlasso)
cash_lasso = all_in(returnlasso)
cash_return_lasso = cash_lasso$cashreturn
cash_overtime_lasso = cash_lasso$overtime

lasso_t  <- get_thres(pred.outlasso,data.val)
IR_lasso_t = lasso_t[1]
returnlasso_t <- lasso_t[-(1:2)]
mean_return_lasso_t = mean(returnlasso_t)
cash_lasso_t = all_in(returnlasso_t)
cash_return_lasso_t = cash_lasso_t$cashreturn
cash_overtime_lasso_t = cash_lasso_t$overtime

output = rbind(output,"lasso" = c(IR_lasso,mean_return_lasso,cash_return_lasso,
                                  IR_lasso_t,mean_return_lasso_t,cash_return_lasso_t))
output

summary(pred.outlasso)
```
LASSO doesn't beat lM1 because it predicts every entry the same value. The major reason is that lasso pick non of the variables as significant and pushed them all to zero. The predicted value is the intercept. It's basically the baseline model. Lasso can give us meaningful results if we set the right random seed, but since we have to try hard to select a specific seed this means the method is not very robust and possibly would not generalize well.

```{r}
plot(cash_overtime_baseline, type='l', xlab="Days", ylab="Cumulative Cash Value ($)")
lines(cash_overtime_lasso)
lines(cash_overtime_lasso_t)
title("Cash Value Over Time, Lasso Models (not different from baseline)")
```


### SVM

```{r svm}
#SVM
library(e1071)
svmM1 <- svm(dnas ~ dbpp2+dcpi1+nas_3d_mean, data=data.in)
svmM2 <- svm(dnas ~ dnas1 + dbpp1 + dcpi1+ dbpp2 + dcpi2, data=data.in)
svmM3 <- svm(dnas ~ .-NASDAQ-return-dbpp-dcpi-CPI-BPP, data=data.in) # all variables
svmM4 <- svm(dnas ~ dnas1+dcpi1+dnas2+
                    nas_3d_mean+nas_3d_max+bpp_3d_median+nas_10d_mean+
                    cumavg.nas,data=data.in) # best for get_thres method
svmM5 <- svm(dnas ~ dnas1+ dbpp1+ dcpi1+ dbpp2+dcpi2+ 
                    nas_3d_mean+nas_3d_max+bpp_3d_median+bpp_3d_max+nas_10d_mean+
                    cumavg.nas,data=data.in) # best for get_quantile method

#predicted return = y_hat
pred.outsvm1   <- predict(svmM1,  newdata=data.val)
pred.outsvm2   <- predict(svmM2,  newdata=data.val)
pred.outsvm3   <- predict(svmM3,  newdata=data.val)
pred.outsvm4   <- predict(svmM4,  newdata=data.val)
pred.outsvm5   <- predict(svmM5,  newdata=data.val)

svm1_q  <- get_quantile(pred.outsvm1,data.val)
IR_svm1 = svm1_q[1]
returnsvm1 <- svm1_q[-(1:3)]
mean_return_svm1 = mean(returnsvm1)
cash_svm1 = all_in(returnsvm1)
cash_return_svm1 = cash_svm1$cashreturn
cash_overtime_svm1 = cash_svm1$overtime

svm1_t  <- get_thres(pred.outsvm1,data.val)
IR_svm1_t = svm1_t[1]
returnsvm1_t <- svm1_t[-(1:3)]
mean_return_svm1_t = mean(returnsvm1_t)
cash_svm1_t = all_in(returnsvm1_t)
cash_return_svm1_t = cash_svm1_t$cashreturn
cash_overtime_svm1_t = cash_svm1_t$overtime


svm2_q  <- get_quantile(pred.outsvm2,data.val)
IR_svm2 = svm2_q[1]
returnsvm2 <- svm2_q[-(1:3)]
mean_return_svm2 = mean(returnsvm2)
cash_svm2 = all_in(returnsvm2)
cash_return_svm2 = cash_svm2$cashreturn
cash_overtime_svm2 = cash_svm2$overtime

svm2_t  <- get_thres(pred.outsvm2,data.val)
IR_svm2_t = svm2_t[1]
returnsvm2_t <- svm2_t[-(1:3)]
mean_return_svm2_t = mean(returnsvm2_t)
cash_svm2_t = all_in(returnsvm2_t)
cash_return_svm2_t = cash_svm2_t$cashreturn
cash_overtime_svm2_t = cash_svm2_t$overtime


svm3_q  <- get_quantile(pred.outsvm3,data.val)
IR_svm3 = svm3_q[1]
returnsvm3 <- svm3_q[-(1:3)]
mean_return_svm3 = mean(returnsvm3)
cash_svm3 = all_in(returnsvm3)
cash_return_svm3 = cash_svm3$cashreturn
cash_overtime_svm3 = cash_svm3$overtime

svm3_t  <- get_thres(pred.outsvm3,data.val)
IR_svm3_t = svm3_t[1]
returnsvm3_t <- svm3_t[-(1:3)]
mean_return_svm3_t = mean(returnsvm3_t)
cash_svm3_t = all_in(returnsvm3_t)
cash_return_svm3_t = cash_svm3_t$cashreturn
cash_overtime_svm3_t = cash_svm3_t$overtime

svm4_q  <- get_quantile(pred.outsvm4,data.val)
IR_svm4 = svm4_q[1]
returnsvm4 <- svm4_q[-(1:3)]
mean_return_svm4 = mean(returnsvm4)
cash_svm4 = all_in(returnsvm4)
cash_return_svm4 = cash_svm4$cashreturn
cash_overtime_svm4 = cash_svm4$overtime


svm4_t  <- get_thres(pred.outsvm4,data.val)
IR_svm4_t = svm4_t[1]
returnsvm4_t <- svm4_t[-(1:2)]
mean_return_svm4_t = mean(returnsvm4_t)
cash_svm4_t = all_in(returnsvm4_t)
cash_return_svm4_t = cash_svm4_t$cashreturn
cash_overtime_svm4_t = cash_svm4_t$overtime

svm5_q  <- get_quantile(pred.outsvm5,data.val)
IR_svm5 = svm5_q[1]
returnsvm5 <- svm5_q[-(1:3)]
mean_return_svm5 = mean(returnsvm5)
cash_svm5 = all_in(returnsvm5)
cash_return_svm5 = cash_svm5$cashreturn
cash_overtime_svm5 = cash_svm5$overtime

svm5_t  <- get_thres(pred.outsvm5,data.val)
IR_svm5_t = svm5_t[1]
returnsvm5_t <- svm5_t[-(1:3)]
mean_return_svm5_t = mean(returnsvm5_t)
cash_svm5_t = all_in(returnsvm5_t)
cash_return_svm5_t = cash_svm5_t$cashreturn
cash_overtime_svm5_t = cash_svm5_t$overtime

output = rbind(output,"svm1" = c(IR_svm1,mean_return_svm1,cash_return_svm1,
                                 IR_svm1_t,mean_return_svm1_t,cash_return_svm1_t),
               "svm2" = c(IR_svm2,mean_return_svm2,cash_return_svm2,
                          IR_svm2_t,mean_return_svm2_t,cash_return_svm2_t),
               "svm3" = c(IR_svm3,mean_return_svm3,cash_return_svm3,
                          IR_svm3_t,mean_return_svm3_t,cash_return_svm3_t),
               "svm4" = c(IR_svm4,mean_return_svm4,cash_return_svm4,
                          IR_svm4_t,mean_return_svm4_t,cash_return_svm4_t),
               "svm5" = c(IR_svm5,mean_return_svm5,cash_return_svm5,
                          IR_svm5_t,mean_return_svm5_t,cash_return_svm5_t))
output

```

The best we can get from svm is svm4_t (SVM 4 with threshold investment strategy), formula =  (dnas1+dcpi1+dnas2+nas_3d_mean+nas_3d_max+bpp_3d_median+nas_10d_mean+cumavg.nas).

```{r}
plot(cash_overtime_svm4_t, type='l', col='green', xlab='Days',ylab='Cumulative Cash Value ($)')
lines(cash_overtime_svm1,col='gray')
lines(cash_overtime_svm2,col='gray')
lines(cash_overtime_svm3,col='gray')
lines(cash_overtime_svm4,col='lightgreen')
lines(cash_overtime_svm5,col='gray')
lines(cash_overtime_svm1_t,col='gray')
lines(cash_overtime_svm2_t,col='gray')
lines(cash_overtime_svm3_t,col='gray')
lines(cash_overtime_svm5_t,col='gray')
lines(cash_overtime_baseline,col='black')
lines(cash_overtime_svm4_t, col='green', lwd=2)
title("Cumulative Cash Value over Time, SVM Models, out-of-sample")
legend(1,1.55e7,legend=c("SVM4 Threshold", "SVM4 Quantile" ,"Baseline"), lty=1,col=c("green","lightgreen", "black"))

```


### Random Forest

```{r randomforest}
library(randomForest)
library(tidyr)

data.in.nona =data.frame(data.in)%>%drop_na()
excl = c('NASDAQ','BPP','CPI','return','dbpp','dcpi')
x.rf.train = data.in.nona[,!(colnames(data.in.nona) %in% excl)]

rf <- randomForest(dnas ~ .,data=x.rf.train,ntree = 200)

pred.outrf = predict(rf, newdata=data.frame(data.val))

rf_q  <- get_quantile(pred.outrf,data.val)
IR_rf = rf_q[1]
returnrf <- rf_q[-(1:3)]
mean_return_rf = mean(returnrf)
cash_rf = all_in(returnrf)
cash_return_rf = cash_rf$cashreturn
cash_overtime_rf = cash_rf$overtime

rf_t  <- get_thres(pred.outrf,data.val)
IR_rf_t = rf_t[1]
returnrf_t <- rf_t[-(1:3)]
mean_return_rf_t = mean(returnrf_t)
cash_rf_t = all_in(returnrf_t)
cash_return_rf_t = cash_rf_t$cashreturn
cash_overtime_rf_t = cash_rf_t$overtime


output = rbind(output,"Random Forest" = c(IR_rf,mean_return_rf,cash_return_rf,
                                          IR_rf_t,mean_return_rf_t,cash_return_rf_t))
output
```
Random Forest doesn't beat linear or svm.

```{r}
plot(cash_overtime_rf,type='l',col='green',xlab='Days',ylab='Cumulative Cash Value ($)')
lines(cash_overtime_rf_t,col='gray')
lines(cash_overtime_rf, col='green')
lines(cash_overtime_baseline,col='black')
title("Cumulative Cash Value over Time, RF Models, out-of-sample")
legend(1,1.37e7,legend=c("RF Quantile" ,"RF Threshold","Baseline"), lty=1,col=c("green","gray", "black"))
```

### XGboost

```{r xgboost}
library(xgboost)
set.seed(1)
xg <- xgboost(data = x.train, label = as.integer(y.train>0), eta = 0.1,max.depth = 25,nrounds = 3, objective = "binary:logistic", verbose = 2)
pred.xg <- predict(xg, x.val)

xg_q  <- get_quantile(pred.xg,data.val)
IR_xg = xg_q[1]
returnxg <- xg_q[-(1:3)]
mean_return_xg = mean(returnxg)
cash_xg = all_in(returnxg)
cash_return_xg = cash_xg$cashreturn
cash_overtime_xg = cash_xg$overtime

xg_t  <- get_thres(pred.xg,data.val)
IR_xg_t = xg_t[1]
returnxg_t <- xg_t[-(1:3)]
mean_return_xg_t = mean(returnxg_t)
cash_xg_t = all_in(returnxg_t)
cash_return_xg_t = cash_xg_t$cashreturn
cash_overtime_xg_t = cash_xg_t$overtime


output = rbind(output,"XGboost" = c(IR_xg,mean_return_xg,cash_return_xg,
                                    IR_xg_t,mean_return_xg_t,cash_return_xg_t))
output
```
xgboost, doesn't beat lm or svm

```{r}
plot(cash_overtime_xg,type='l',col='green',xlab='Days',ylab='Cumulative Cash Value ($)')
lines(cash_overtime_xg_t,col='gray')
lines(cash_overtime_xg, col='green')
lines(cash_overtime_baseline,col='black')
title("Cumulative Cash Value over Time, XGBoost Models, out-of-sample")
legend(1,1.34e7,legend=c("XGBoost Quantile" ,"XGBoost Threshold","Baseline"), lty=1,col=c("green","gray", "black"))
```

Let's look at each method:
```{r}
output[1:5,]
```
Among OLS, we pick lm5(Quantile Method) as our best performing model.

```{r}
output[7:11,]
```
Among SVM, we pick svm4(Threshold method) as our best performing model.

Let's see final comparison
```{r}
final = data.frame(InformationRatio = c(IR_baseline,IR_lM5,IR_lasso,IR_svm4_t,IR_rf,IR_xg),
           MeanReturn = c(mean_return_baseline,mean_return_lM5,mean_return_lasso,
                          mean_return_svm4_t,mean_return_rf,mean_return_xg),
           CashReturn = c(cash_return_baseline,cash_return_lM5,cash_return_lasso,
                          cash_return_svm4_t,cash_return_rf,cash_return_xg))
rownames(final) <-c("baseline","OLS5_quantile","lasso_quantile","SVM4_threshold","RandomForest_quantile","xgboost_quantile")
final[order(-final$InformationRatio),]
```


```{r}
plot(cash_overtime_svm4_t,type='l',col='green',xlab='Days',ylab='Cumulative Cash Value ($)')
lines(cash_overtime_svm4_t,col='green')
lines(cash_overtime_lM5, col='blue')
lines(cash_overtime_xg, col='orange')
lines(cash_overtime_rf, col='purple')
lines(cash_overtime_lasso, col='gray')
lines(cash_overtime_baseline, col='black')
lines(cash_overtime_svm4_t,col='green')
title("Cumulative Cash Value over Time, All Models, out-of-sample")
legend(1,1.55e7,legend=c("SVM4 Threshold" ,"OLS5 Quantile","RF Quantile","XGBoost Quantile","Baseline"), lty=1,col=c("green","blue", "purple","orange","black"))
```


```{r}
data.frame(Model = c("Best SVM","Lasso=Baseline","XGBoost","Best OLS","Random Forest"),
           OutOfSampleRSquared = format(c( 
1-(sum((data.val$dnas - pred.outsvm4)^2)/sum((data.val$dnas -mean(data.in$dnas[!is.na(data.in$dnas)]))^2)), 
1-(sum((data.val$dnas - pred.outlasso)^2)/sum((data.val$dnas - mean(data.in$dnas[!is.na(data.in$dnas)]))^2)),
1-(sum((data.val$dnas - pred.xg)^2)/sum((data.val$dnas - mean(data.in$dnas[!is.na(data.in$dnas)]))^2)),
1-(sum((data.val$dnas - pred.outl5)^2)/sum((data.val$dnas - mean(data.in$dnas[!is.na(data.in$dnas)]))^2)),
1-(sum((data.val$dnas - pred.outrf)^2)/sum((data.val$dnas - mean(data.in$dnas[!is.na(data.in$dnas)]))^2))
),scientific=F)
)
```

```{r}
barplot(c(
mean(sign(data.val$dnas) == sign(pred.outsvm4)),
mean(sign(data.val$dnas) == sign(pred.xg)),
mean(sign(data.val$dnas) == sign(pred.outlasso)),
mean(sign(data.val$dnas) == sign(pred.outl5)),
mean(sign(data.val$dnas) == sign(pred.outrf))), 
names.arg=c("Best SVM", "XGBoost","Lasso=Baseline","Best OLS","Random Forest"),
col=c("green","orange","gray","blue","purple"))
title("Proportion of times dnas sign was correctly predicted")

```


## Summary
the best performing models are SVM4_threshold and OLS5_quantile.
```{r}
svmM4 <- svm(dnas ~ dnas1+dcpi1+dnas2+
                    nas_3d_mean+nas_3d_max+bpp_3d_median+nas_10d_mean+
                    cumavg.nas,data=data.in) # best for get_thres method

pred.outsvm4   <- predict(svmM4,  newdata=data.val)

svm4_t  <- get_thres(pred.outsvm4,data.val)
threshold_svm4 = svm4_t[2]


print(paste("SVM model: we invest in all predicted value of dnas that's >=",threshold_svm4))
```


```{r}
lM5     <- lm(dnas ~ dnas1+bpp_3d_median,data=data.in)
pred.outl5   <- predict(lM5, newdata=data.val)
lM5_q  <- get_quantile(pred.outl5,data.val)
q1 = lM5_q[2]
q2 = lM5_q[3]

print(paste("OLS model: we invest in all predicted value of dnas that's >=",q1,"and <=",q2))
```

## Final in-sample metrics:


```{r}
insampledata = data[complete.cases(data),]

lM5     <- lm(dnas ~ dnas1+bpp_3d_median,data=insampledata)
pred.wholesample  <- predict(lM5, newdata=insampledata)

#lM5_q  <- get_quantile(pred.wholesample,insampledata)

buy   <- c(as.numeric((pred.wholesample>=2.66644464508991)&
                                (pred.wholesample<=3.90782832612255)))

dailyreturn  <- (insampledata$return-1)*buy

IRlm5  <- mean(dailyreturn)/sd(dailyreturn)
print(IRlm5)
      
mean_return_lM5 = mean(dailyreturn)
print(mean_return_lM5)

cash_lm5 = all_in(dailyreturn)
cash_return_lm5 = cash_lm5$cashreturn
cash_overtime_lm5 = cash_lm5$overtime
print(cash_return_lm5)

```


```{r}
svmM4 <- svm(dnas ~ dnas1+dcpi1+dnas2+
                    nas_3d_mean+nas_3d_max+bpp_3d_median+nas_10d_mean+
                    cumavg.nas,data=insampledata)

pred.wholesample   <- predict(svmM4,  newdata=insampledata)

buy   <- c(as.numeric(pred.wholesample>=6.5))

dailyreturn  <- (insampledata$return-1)*buy

IRsvm  <- mean(dailyreturn)/sd(dailyreturn)
print(IRsvm)
      
mean_return_svm4 = mean(dailyreturn)
print(mean_return_svm4)

cash_SVM4 = all_in(dailyreturn)
cash_return_SVM4 = cash_SVM4$cashreturn
cash_overtime_SVM4 = cash_SVM4$overtime
print(cash_return_SVM4)
```

```{r}
buy_baseline   <- 1

dailyreturn  <- (insampledata$return-1)*buy_baseline

IRbaseline  <- mean(dailyreturn)/sd(dailyreturn)
print(IRbaseline)
      
mean_baseline = mean(dailyreturn)
print(mean_baseline)

cash_bline = all_in(dailyreturn)
cash_return_bline = cash_bline$cashreturn
cash_overtime_bline = cash_bline$overtime
print(cash_return_bline)
```

```{r}
data.frame(Model=c("SVM4Threshold","OLS5Quantile","Baseline"), 
           InformationRatio=c(IRsvm,IRlm5,IRbaseline),
           MeanReturn=c(mean_return_svm4,mean_return_lM5,mean_baseline),
           CashReturn = c(cash_return_SVM4,cash_return_lm5,cash_return_bline))
```



```{r}
plot(cash_overtime_SVM4,type='l',col='green',xlab='Days',ylab='Cumulative Cash Value ($)')
lines((cash_overtime_lm5),col='blue')
lines((cash_overtime_bline),col='black')
title("Cumulative Cash Value over Time, Best Models, in-sample")
legend(1,4.5e7,legend=c("SVM4 Threshold" ,"OLS5 Quantile","Baseline"), lty=1,col=c("green","blue","black"))
```


### Extra: try short selling, does not work better
```{r}
svmM4 <- svm(dnas ~ dnas1+dcpi1+dnas2+
                    nas_3d_mean+nas_3d_max+bpp_3d_median+nas_10d_mean+
                    cumavg.nas,data=data.in)

pred.wholesample   <- predict(svmM4,  newdata=insampledata)

buy   <- (as.numeric(pred.wholesample>=6.5))

buy[pred.wholesample<-6.5]   <- -1

dailyreturn  <- (insampledata$return-1)*buy

IR  <- mean(dailyreturn)/sd(dailyreturn)
print(IR)
      
mean_return_svm4 = mean(dailyreturn)
print(mean_return_svm4)

cash_return_svm4 = all_in(dailyreturn)$cashreturn
print(cash_return_svm4)
```






